---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# FuzzyFishHS

<!-- badges: start -->
<!-- badges: end -->

FuzzyFishHS provides functions to develop and inspect zero-order presence-absence Takagi–Sugeno–Kang fuzzy rule-based systems for fish habitat evaluation within environmental flow assessemnt studies. 

## Installation

You can install the development version of FuzzyFishHS from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("RafaMMas/FuzzyFishHS")
```

## Example

### Model development

```{r example}
library(FuzzyFishHS)
```

Create the Fuzzy Rule-Based System (FRBS) using the function FRBS. This function simply lists the following five arguments:

- *ImpVariables* the names of the input variables in vector format
- *Range* a matrix with as many columns as input variables with the minimum and maximum expected values for each variable
- *MFfunction* the selected membership function
- *MFparameters* the parameters of the membership functions. It should be a matrix with as many columns as fuzzy sets. The columns for each variable must include the variable name as indicated in *ImpVariables*.
- *Consequents* the zero-order TSK fuzzy rule consequents.

You can choose among the 10 different membership functions implemented in the packages. The package includes a function (*HILL.CLIMB.FRBS.Super.Fast.TSS*) to train (optimise) the fuzzy rules' consequents. At this point you can add a random vector with as many consequents as fuzzy rules. 
<br>
```{r create FRBS, echo = TRUE}
Example.FRBS <- Create.FRBS.R(ImpVariables = c("Velocity", "Depth", "Substrate.index", "Cover.index"),
                       Range = matrix(c(0, 1.164, 0, 2.67, 0, 8, 0, 1), ncol = 4, dimnames = list(NULL, c("Velocity", "Depth", "Substrate.index", "Cover.index")
                                                                                                  )),
                       MFfunction = PIMF,
                       MFparameters = matrix(c(0, 0, 0.0377, 0.4, 0.0377, 0.4, 1.164, 1.164, 0, 
                                               0, 0.46, 0.8767, 0.46, 0.8767, 1.3067, 2.67, 1.3067, 2.67, 2.67, 
                                               2.67, 0, 0, 0, 0.5333, 0, 0.5333, 3.8667, 6, 3.8667, 6, 8, 8, 
                                               0, 0, 0, 1, 0, 1, 1, 1), ncol = 10, 
                                           dimnames = list(NULL, c("VelocityL", "VelocityH", "DepthL", 
                                                                   "DepthM", "DepthH", "Substrate.indexL", "Substrate.indexM", "Substrate.indexH", 
                                                                   "Cover.indexL", "Cover.indexH"))),
                       Consequents = c(0, 0, 1, 0, 1, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 1, 1, 0, 0.5, 0, 0, 1, 0, 0, 1, 1, 0, 0, 0, 0, 0)
                       )
```
<br>
Once the *FRBS* has been created it is recommended to plot the membership functions and check whether they are correct.
<br>
```{r plot membership functions, fig.dim=c(10, 4), dpi=600, echo = TRUE}
PlotMF(FRBS = Example.FRBS, n.pt = 9999,
       data = NULL,
       Title = "Lepomis gibbosus")
```
<br>
<br>
The function *PlotMF* allows including the training dataset to compare the membership functions with the distribution of the input variables per output class (*i.e.,* presence or absence). Pink corresponds to presence data and grey to absence.
<br>
```{r plot membership functions and data, fig.dim=c(10, 4), dpi=600, echo = TRUE}
PlotMF(FRBS = Example.FRBS, n.pt = 9999,
      data = Lepomis.gibbosus.dataset,
      Title = "Lepomis gibbosus")
```

### Fuzzy rules training

In order to boost rules' training the membership of each datum to each fuzzy rule is calculated beforehand. It can be done employing the function *FUZZIFY.FRBS.Fast*.
<br>
```{r obtain membership, echo = TRUE}
Memberships <- FUZZIFY.FRBS.Fast(Data = Lepomis.gibbosus.dataset[,Example.FRBS$ImpVariables],
                                 FRBS = Example.FRBS)
```
<br>
Once the membership is calculated it it fed to the function *HILL.CLIMB.FRBS.Super.Fast.TSS*. The hill climbing algorithm starts from a random set of consequents and modifies one consequance at a time retaining those changes tha improve model performance (*i.e.,* the True Skill Statistic). Threfore, we recommend repeating the optimisation multiple times to avoid getting traped in local minima.
<br>
```{r rules training, echo = TRUE}
Optimised.consequents <- HILL.CLIMB.FRBS.Super.Fast.TSS(Membership = Memberships,
                                                        Training.dataset = Lepomis.gibbosus.dataset,
                                                        Trials = 9)
print(Optimised.consequents)

```
<br>

After the hill climbing algorithm finished, the random consequents can be substitutted by those optimal and predictions can be done as follow:

```{r par 1, echo = FALSE}

op <- par(oma = c(0.5, 0.5, 0.5, 0.5), mar=c(5, 5.5, 0.5, 0.5), cex.axis = 1.5, cex.lab = 2, mgp = c(2, 1, 0))

```

```{r prediction, out.height="50%", fig.align = 'center', out.width="40%", dpi=300, echo = TRUE}

Predict <- PREDICT.FRBS.Fast(Data = Lepomis.gibbosus.dataset[,Lepomis.gibbosus.FRBS$ImpVariables],
                             FRBS = Example.FRBS)

op <- par(oma = c(0.5, 0.5, 0.5, 0.5), mar=c(5, 6, 0.5, 0.5), cex.axis = 1.5, cex.lab = 2, mgp = c(3.5, 1, 0))

plot(Lepomis.gibbosus.dataset$Species, Predict[2,],
                              bty = "n", las = 1, col = "orangered", xlab = "Observed", ylab = "Predicted")
abline(lm(Predict[2,] ~ Lepomis.gibbosus.dataset$Species), col = "dodgerblue")

par(op)

```

```{r par 2, echo = FALSE}

par(op)

```

<br>

